---
title: "Homework 8"
author: "Anh Dao"
date: "`r Sys.Date()`"
echo: false
format:
  gfm:
    toc: true
    html-math-method: webtex
editor: 
  markdown: 
    wrap: 72
---

```{r}
#| message: false
#| warning: false
library(tidyverse)
library(brms)
library(posterior)
library(readr)
library(bayesplot)
library(modelsummary)
library(dplyr)
options(brms.backend = "cmdstanr", mc.cores = 2)
```

# Research question:

Does reward consumption increase throughout CBT and ACT group therapy?

The current study aims to assess group and individual trajectories of
change of reward consumption following cognitive behavioral therapy
(Beck, 1979) and acceptance and commitment therapy (ACT; Hayes et al.,
1999).

# Variables

-   'rr1': Reward consumption "I found it enjoyable/fulfilling to do
    things."

    -   Scale: 1= Not at All; 2 = A Little; 3 = Moderately; 4 = Quite a
        Bit; 5 = Extremely

    -   Note that this variable is ordinal, but the model cannot run if
        we set ORDINAL = RR1. The distribution of responses does not
        show ceiling/floor effects, and based on the skewness and
        kurtosis, it may be safe to run the analysis with the assumption
        that RR1 is continuous.

-   'time': Assessment instance

-   'Beta1': Autoregressive effect of reward consumption

-   'trend1': Effect of time on reward consumption

-   'logV1': Log variance of reward consumption

-   '\_v': Represents variance of each parameter

## Data Import

```{r}
#| message: false

#Read in data 
results <- read_csv("rewbayes573.csv")
rewdata <- read_csv("data_mech.csv")

#Save as draws
draws_x <- as_draws(results)
draws_x <- as.data.frame(draws_x)
```

## Variable Summary

Table @tbl-summ-var shows the summary statistics of reward consumption.

```{r}
#| label: tbl-summ-var
#| message: false
library(modelsummary)
datasummary(rr1  * (N + Mean + SD + Min + Max + Histogram)~ 1, data = rewdata)
```

Model

Let L be the intercept, D the autoregressive effect, Time the time trend, and e the residual variance.

Within:
$$
  \begin{aligned}
    Rew_{tj} = l_{Rew_{j}} + D_{Rew_{j}}Time_{tj} + e_{Rew_{tj}} \\
    e_{Rew_{tj}} = \beta_{Rew_{j}}e_{Rew_{(t-1)}} + \delta_{Rew_{tj}}\\
  \end{aligned}
$$
$$
  \begin{aligned}
    D_{Rew_{j}} =  D_{Rew} + \mu_{j} \\
  \end{aligned}
$$
Prior: 
$$
  \begin{aligned}
    \ l_{Rew_{j}} & \sim N(0.2, 0.05) \\
    \ D_{Rew_{j}} & \sim \ N(0.1, 0.01) \\
    \ Time_{tj} & \sim \Gamma^{-1}(-1, 0.50) \\
    \ e_{Rew_{tj}}  & \sim \Gamma^{1}(1.5, 0.20) \\
  \end{aligned}
$$ 

Analysis 

We used 2 chains, each with 20,000 iterations (first 10,000 as
warm-ups).

## Mplus code

```         
TITLE:  RIPTIDE

DATA:
  FILE IS "C:\Users\anhdao\OneDrive - Vanderbilt\Desktop\RIPTIDE Data\
  data_mech_mplus.csv";

VARIABLE:
  NAMES ARE record group no  instance qol_td pro_phq cr1 cr2
    cr3 ba1 ba2 ba3 ps acc mind sac def val ca dist supp avoid rum ss cc1 cc2   
    cc3 rr1 rr2 pf1 pf2 pf3 er aveact actskill se affect1 affect2
    affect3 affect4 affect5 affect6 affect7 affect8 affect9 affect10
    na pa phq1 phq2 phq3 phq4 phq5 phq6 phq7 phq8 phq9; 
  USEVARIABLES ARE record rr1 time;
  WITHIN = time;
  MISSING IS .;
  CLUSTER IS record; 
  LAGGED=rr1(1);
  TINTERVAL = no(1); 
DEFINE: 
  time=no;
  
ANALYSIS:
  TYPE = TWOLEVEL RANDOM; 
  ESTIMATOR = BAYES; 
  BITERATIONS = (20000); 
  PROCESSORS =2;
  CONVERGENCE = 0.005;

MODEL:
  %WITHIN%  
    BETA1  | rr1^ ON rr1^1;  
    trend1 | rr1 ON time; 
    LogV1 | rr1;           

   %BETWEEN%
   [rr1]; 
   [BETA1];
   [LogV1];
   [trend1]; 
   
   rr1; 
   BETA1; 
   LogV1; 
   trend1;

OUTPUT: 
  tech1 tech8 cinterval(hpd) standardized (cluster);
SAVEDATA:
  SAVE = FSCORES(200);
  PREDICTORS = ytilde;        ! Save posterior predictive values (simulated values)
  FILE = rew573_pred.dat; 
  PLOT: TYPE = PLOT3 DRIFT (0, 5, 0.1);
```

# Results

```{r}
num_draws <-  20000
num_chains <- 2

# Total iterations and number of warm-ups per chain
num_draws_per_chain <- num_draws / num_chains
num_warmup <- num_draws_per_chain 

# Discard warm-up iterations for each chain
used_draws <- draws_x[draws_x$.iteration > num_warmup, ]
```

@tbl-summ shows the posterior distributions.

```{r}
#| label: tbl-summ
#| tbl-cap: Posterior summary of the model parameters.

#Select the main effect
#trend1_draws <- subset_draws(used_draws, variable = "trend1")
#trend1_draws <- as_draws_array(trend1_draws)
#summary_trend1 <- summarize_draws(trend1_draws)
#print(summary_trend1)

summ <- summarise_draws(used_draws)
knitr::kable(summ, digits = 2)
mcmc_dens(used_draws)
```

As shown in the rank histogram in @fig-rank-hist-fit below, the chains
mixed well.

```{r}
#| label: fig-rank-hist-fit
#| fig-cap: Rank histogram of the posterior distributions of model parameters.

# Trace plot
mcmc_trace(used_draws)
```

The analysis showed that on average, group participants' reward consumption did not change significantly over time with a posterior mean of `r round(summ$mean[5], 2)` and a 90% CI of [`r round(summ$q5[5], 2)`, `r round(summ$q95[5], 2)`].